---
title: "3. Introduction to hypothesis testing via binomial tests"
bibliography: ../references.bib
editor: 
  markdown: 
    wrap: 72
---

<!-- COMMENT NOT SHOW IN ANY OUTPUT: Code chunk below sets overall defaults for .qmd file; these inlcude showing output by default and looking for files relative to .Rpoj file, not .qmd file, which makes putting filesin different folders easier  -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Remember you should

-   add code chunks by clicking the *Insert Chunk* button on the toolbar
    or by pressing *Ctrl+Alt+I* to answer the questions!
-   **knit** your file to produce a markdown version that you can see!
-   save your work often
    -   **commit** it via git!
    -   **push** updates to github

## Overview

This practice reviews the [Hypothesis testing starting with binomial
tests lecture](../chapters/Binomial.qmd).

## Hypothesis Testing and the Binomial Distribution

### Example

Using the bat paper from class (Geipel et al. 2021), let's consider how
to analyze data showing all 10 bats chose the walking over the
motionless model.

```{r}
binom.test(10,10)
```

We use the binom.test function. We only need arguments for \# of
succeses and \# of trials. By default it runs a 2-sided test against a
null hypothesis value of p = .5. You can see how to update thee options
by looking at the help file.

```{r, eval=F}
?binom.test
```

Note the confidence interval is assymetric since its estimated to be 1!
We can see other options using the binom.confint function from the
*binom* package.

```{r}
library(binom)
binom.confint(10,10)
```

All of these correct for the fact that most intervals use a normal
approximation, which as you remember from our earlier discussions is not
good when sample sizes are small and/or the p parameter is extreme
(close to 0 or 1).

## Practice!

Make sure you are comfortable with null and alternative hypotheses for
all examples.

### 1

Are people eared (do they prefer one ear or another)? Of 25 people
observed while in conversation in a nightclub, 19 turned their right ear
to the speaker and 6 turn their left ear to the speaker. How strong is
the evidence for eared-ness given this data (adapted from Analysis of
Biological Data)?

\* state a null and alternative hypothesis

```{r}
#Null Hypothesis: People do not have a preference in which ear they use to hear making the probability .5 for either ear.
#Alternative: People do have a preference of which ear they use more commonly.
```

\* calculate a test statistic (signal) for this data

```{r}
19/25 #using the data to work with the right ear
```

\* Make you understand how to construct a null distribution + using
sampling/simulation (code or written explanation) + by using an
appropriate distribution (code or written explanation)

```{r}
sampling_experiment = rbinom(10000, 25, .5)
hist(sampling_experiment, breaks = 0:25, xlab = "# of Right-eared people out of 25", ylab = "Probability of being drawn \n from population of p = 0.5", cex.main = 2, cex.axis = 1.5, cex.lab = 2)
```

```{r}
using_distribution = dbinom(0:25,25,.5)
using_distribution

sum(using_distribution)
```

```{r}
Number_righteared = c(0:25)
pdf = data.frame(Number_righteared, using_distribution)
plot(0:25, using_distribution)
```

\* Calculate and compare p-values obtained using + simulation
(calculation won’t be required on test, but make sure you understand!)
(code or written explanation) + equations for binomial distribution
(code or written explanation) + R functions (required)(code)

```{r}
#for the r code section
binom.test(19,25,0.5)
```

\* Calculate a 95% confidence interval for the proportion of people who
are right-eared

```{r}
binom.confint(19, 25, conf.level= 0.95, methods = "all")
```

\* How do your 95% confidence interval and hypothesis test compare?

The confidence interval test gave me a closer interval of .56-.89
whereas the binomial test gave a larger yet similar interval.
Regardless, the p-values were below the alpha level of 0.05 so the null
hypothesis is rejected that people do not have an ear preference.

### 2

A professor lets his dog take every multiple-choice test to see how it
compares to his students (I know someone who did this). Unfortunately,
the professor believes undergraduates in the class tricked him by
helping the dog do better on a test. It’s a 100 question test, and every
questions has 4 answer choices. For the last test, the dog picked 33
questions correctly. How likely is this to happen, and is there evidence
the students helped the dog?

**MAKE SURE TO THINK ABOUT YOUR TEST OPTIONS**

```{r}
binom.test(33,100,p=.25)
```

considering the p-value we fail to reject the hypothesis that the
students did not help the dog. so they might have helped him:). This has
a p-value of 0.08222 so it is about 8% likely that they helped him. From
the given information, I don't see evidence to conclude they helped the
dog; It is about 92% possible that they did not help him because the dog
had a 1/4 chance of choosing the right answer for each question.
